### 2019-09-27 LaTeX-Master-Vorlage/main.tex
working working doing stuff

\include{settings}

%%Titles - Uncomment one section of titles

%%Used for titleGraduation
\include{titles/titleGraduation}
\logoPathL{} %just leave empty to hide logo
\logoWidthL{0.5}
\logoPathR{resources/uni_bern_logo.jpg} %just leave empty to hide logo
\logoWidthR{1}
\gradeType{Bachelor of Science (B.Sc.)}
\secondExaminer{Max Mustermann}

%%Used for titleResearchPaper
%\include{titles/titleResearchPaper}
%\subTitle{Ein optionaler Untertitel der Arbeit}
%\researchPart{A}

%%Used by all titles
\title{Stable Neo-Hookean Flesh Simulation}
\author{Corina Danja Masanti}
\matrikelnr{15-128-655} % just leave empty to hide number
\submitDate{01.09.2019}
\firstExaminer{Prof. Dr. David Bommes}
%%End Titles

\makeindex[title=Stichwortverzeichnis, options=-s indexstyle.ist, intoc]
\indexsetup{level=\chapter*,toclevel=chapter}

\makeglossaries
\loadglsentries{glossary_and_acronyms.tex}
\setacronymstyle{long-short}

\begin{document}

\pagenumbering{alph} %fix for same identifier warning, character is not show in title
\maketitle

\pagenumbering{Roman}

\include{chapter/Vorwort} \clearpage
\include{chapter/Abstract} \clearpage

\tableofcontents \newpage

\pagenumbering{arabic}
\input{chapter/Einleitung} \clearpage
\input{chapter/Background} \clearpage
%\input{chapter/Beispiele} \clearpage
\input{chapter/Paper} \clearpage

\pagenumbering{Alph}
\listoffigures \clearpage
\listoftables \clearpage
\lstlistoflistings \clearpage

\printindex \clearpage

\printglossary[title={Glossar}] \clearpage
\printglossary[style=dottedlocations,type=\acronymtype,title={Abkürzungsverzeichnis}] \clearpage

\printbibliography[heading=bibintoc, keyword={book}, title={Literature}]\clearpage
\printbibliography[heading=bibintoc, keyword={article}, title={Articles}]\clearpage
\printbibliography[heading=bibintoc, keyword={online}, title={Online Resources}]\clearpage
\printbibliography[heading=bibintoc, type=misc,title={Various References}] \clearpage
\printbibliography[heading=bibintoc, keyword={image}, title={Images Resources}]\clearpage

% Anhang
\input{chapter/Anhang}

% Eigenständigkeitserklärung
\input{chapter/Eigenstaendigkeitserklaerung}

\end{document}