### 2019-09-30 LaTeX-Master-Vorlage/chapter/Introduction.tex
eigenstaendigkeitserklaerung, table of contents ueberarbeitet

\chapter{Introduction}
\section{Motivation}
The goal of this work is for a regular computer science student to give the necessary physical and mathematical background to understand the field of animation.

\section{Structure}
I will start with the background and then go on to the actual paper.