### 2020-02-24 LaTeX-Master-Vorlage/glossary_and_acronyms.tex

\newglossaryentry{latex}
{
        name=latex,
        description={Is a mark up language specially suited for 
scientific documents}
}

\newacronym{gcd}{GCD}{Greatest Common Divisor}
\newacronym{svd}{SVD}{Singular Value Decomposition}