### 2021-10-19 ba-thesis-masanti/main.tex

\include{settings}

%%Titles - Uncomment one section of titles

%%Used for titleGraduation
\include{titles/titleGraduation}
\logoPathL{} %just leave empty to hide logo
\logoWidthL{0.5}
\logoPathR{resources/uni_bern_logo.jpg} %just leave empty to hide logo
\logoWidthR{1}
\gradeType{Bachelor of Science (B.Sc.)}
\secondExaminer{Max Mustermann}

%%Used for titleResearchPaper
%\include{titles/titleResearchPaper}
%\subTitle{Ein optionaler Untertitel der Arbeit}
%\researchPart{A}

%%Used by all titles
\title{Flesh Simulation with Application to Character Animation}
\author{Corina Danja Masanti}
\matrikelnr{15-128-655} % just leave empty to hide number
\submitDate{06.11.2020}
\firstExaminer{Prof. Dr. David Bommes}
%%End Titles

\makeindex[title=Stichwortverzeichnis, options=-s indexstyle.ist, intoc]
\indexsetup{level=\chapter*,toclevel=chapter}

\makeglossaries
\loadglsentries{glossary_and_acronyms.tex}
\setacronymstyle{long-short}

\begin{document}

\pagenumbering{alph} %fix for same identifier warning, character is not show in title
\maketitle

\pagenumbering{Roman}

%\include{chapter/Vorwort} \clearpage
\include{chapter/Abstract} \clearpage
\include{chapter/Acknowledgements} \clearpage

\tableofcontents \newpage

\pagenumbering{arabic}
\input{chapter/Introduction} \clearpage
\input{chapter/Background} \clearpage
%\input{chapter/Beispiele} \clearpage
\input{chapter/Paper} \clearpage
\input{chapter/Experiments} \clearpage

\pagenumbering{Alph}
\listoffigures \clearpage
\listoftables \clearpage

\lstlistoflistings \clearpage

\printindex \clearpage

%\printglossary[title={Glossary}] \clearpage

\printglossary[style=dottedlocations,type=\acronymtype,title={List of abbreviations}] \clearpage

% trying to make bibliography work
\defbibfilter{booksAndPapers}{
  type=article or
  type=inproceedings or
  type=incollection or
  type=book or
  type=inbook or
  type=misc
}
\defbibfilter{onlineSource}{
  type=online
}

\printbibliography[heading=bibintoc, filter=booksAndPapers, title={Bibliography}]\clearpage
\printbibliography[heading=bibintoc, filter=onlineSource, title={Online Sources}]\clearpage
\printbibliography[heading=bibintoc, keyword={image}, title={Figure Sources}]\clearpage

%\printbibliography[heading=bibintoc, keyword={book}, title={Literature}]\clearpage
%\printbibliography[heading=bibintoc, keyword={article}, title={Articles}]\clearpage
%\printbibliography[heading=bibintoc, keyword={online}, title={Online Resources}]\clearpage
%\printbibliography[heading=bibintoc, type=misc,title={Various References}] \clearpage
%\printbibliography[heading=bibintoc, keyword={image}, title={Images Resources}]\clearpage

% Anhang
%\input{chapter/Anhang}

% Eigenständigkeitserklärung
%\input{chapter/Declaration_of_Originality}
% This is required since 2012!!
\includepdf{Erklaerung.pdf}

\end{document}